### 2023-04-09 main/chapter_experiment.tex
write about magnets

Radiofrequency (RF) cavities are the technology used to accelerate protons in
the LHC.
They operate by setting up electron beams to oscillate at a frequency that
resonates with a specially designed cavity, this producing strong oscillating
electromagnetic fields.
These electromagnetic fields oscillate at $400\,\mathrm{MHz}$ in the LHC, and
must be carefully tuned jointly with the incident proton beams to align in
their timing.
When perfectly timed, the RF cavity will accelerate protons that are less
energetic that the target beam energy, and decelerate any that are more
energetic.
This achieves a dynamically stable acceleration.
The LHC uses $16$ superconducting RF cavities contained in four separate
cryostatic modules~\cite{lhcDesignreport2004}.

% TODO magnets
% - dipole for steering
% - quadrupole for focussing (pairs)
% - higher orders why? septupole at beam entry
% TODO size of beam
% TODO luminosity and its uncertainty
% TODO pile-up
\TODO{Distribution in the mean number of primary vertices per crossing, where the
variability in that mean arises from differing beam conditions over run-time.}
Figure~\ref{fig:experiment_run_2_mu}

>>>>

Radiofrequency (RF) cavities are the technology used to accelerate protons in
the LHC.
They operate by setting up electron beams to oscillate at a frequency that
resonates with a specially designed cavity, this producing strong oscillating
electromagnetic fields.
These electromagnetic fields oscillate at $400\,\mathrm{MHz}$ in the LHC, and
must be carefully tuned jointly with the incident proton beams to align in
their timing.
When perfectly timed, the RF cavity will accelerate protons that are less
energetic that the target beam energy, and decelerate any that are more
energetic.
This achieves a dynamically stable acceleration.
The LHC uses sixteen superconducting RF cavities, which are contained in four
separate cryostatic modules~\cite{lhcDesignreport2004}.

% Magnets
Magnets are used to steer and focus the LHC's proton beams.
% - dipole for steering
Dipole magnets conduct steering by creating constant magnetic fields, with
north and south magnetic poles on diametrically-opposite sides of the beam.
These dipole fields steer the beam by applying constant Lorentz forces
perpendicular to the beam's motion.
% - quadrupole for focussing (pairs)
Quadrupole magnets focus the beam to counteract its natural dispersion, with
alternating magnetic poles placed after each $90^\circ$ rotation around the
beam.
A quadrupole magnet is focussing in one axis, and defocussing in a
perpendicular axis.
However, alternating pairs of quadrupole magnets can achieve
a total focussing effect; their effects are not linear.

% higher orders
Sextupole, octupole, and decapole magnets are used to implement detailed
corrections with each of the dipoles magnets.
Chromatic aberrations in optics are errors that arise from the colour-dependent
behaviour of light.
The analogous momentum-dependent behaviour in particle beams is known as
`chromaticity'.
Sextupole magnets are used in the LHC to correct for chromaticity at each
quadrupole~\cite{Guiducci:1992kb, lhc2008machine}.

% TODO size of beam
% TODO luminosity and its uncertainty
% TODO pile-up
\TODO{Distribution in the mean number of primary vertices per crossing, where the
variability in that mean arises from differing beam conditions over run-time.}
Figure~\ref{fig:experiment_run_2_mu}
